### 2023-07-07 src/manuscript.qmd
another draft'

---
title: |
  Economic concerns appear to be weak predictors of white political identity
short-title: Economic concerns
published: "Manuscript draft: please do not share without author permission."
code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/white-identity-sources>."
author:
  - name: Damon C. Roberts
    email: damon.roberts-1@colorado.edu
    orcid: 0000-0002-4360-3675
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
abstract: |
    Do economic or political threats explain reported white identity? Overall, the social identity literature would suggest that white identity would increase in response to economic threats. However, a number of those that study white identity, specifically, argue that it results from concern about political influence. Considering what whiteness means historically and contemporaneously, I argue that we should expect that political threats reflect stronger associations with white identity. Using data from the 2012, 2016, and 2020 American National Election Study, I consider a single model using penalized regression containing proxies of economic and political threats. I find evidence suggesting that in the post-Trump era, white political identity is strongly associated with reported feelings of Whites’ loss of political influence as opposed to economic threats, as some suggest and may expect.
thanks: |
    Word count: 3575. I would like to thank Jennifer Wolak for her feedback on the many drafts of this manuscript. I would also like to thank the Keller Center for the First Amendment for the research appointment that gave me the space to work on the early stages of this project. I would also like to thank Anand Sokhey and John Griffin for their advice and suggestions in the early drafts of this paper. I would also like to thank Tabitha Bonilla, Nicole Yadon, and participants and audience members of a Midwestern Political Science Association panel for which an earlier draft this idea was presented. I would also like to thank Tamar Malloy, Andrew Philips, Kathryn Schauer, Rachel O'Neal, Alexander Jensen, Komal Kaur, Andrew Engelhardt, and Madeline Mader for their encouragement and feedback at various stages of this project's development. All errors contained in this manuscript are ultimately mine.
additional-info: |
  I have no known conflict of interest to disclose.
keywords:
  - white identity
  - economic threat
  - cultural threat
date: today
bibliography: "../assets/wid.bib"
format:
  hikmah-pdf:
    cite-method: citeproc
    latex-output-dir: ../out/
    filters:
    - "_extensions/andrewheiss/wordcount/citeproc.lua"
    - "_extensions/andrewheiss/wordcount/wordcount.lua"
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
    linestretch: 2
    fontsize: 12pt
    appendix-style: none
    include-in-header:
      text: |
        \usepackage{pdflscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
linestretch: 2
execute:
  echo: false
  warning: false
  message: false
params:
  replicate: false
---

```{r}
#| label: block-setup
# set seed
set.seed(12102022)
# modularly load functions
box::use(
  dplyr = dplyr[
    mutate
    , filter
    , case_when
    , select
  ]
  , tibble = tibble[
    tribble
  ]
  , modelsummary = modelsummary[
    datasummary_skim
    , modelsummary
  ]
  , parameters = parameters[
    parameters
    , standardize_names
  ]
  , ggplot2 = ggplot2[
    ggplot
    , aes
    , position_dodge
    , theme_minimal
    , geom_vline
    , labs
    , theme
    , element_text
    , scale_y_discrete
  ]
  , tidybayes = tidybayes[
    geom_pointinterval
  ]
  , data.table[
    setnames
  ]
)
# If doing a full replication, change replicate option in yaml section
if (params$replicate == TRUE) {
  source("./models/model_fitting.R")
}
```

```{r}
#| label: load-fitted-model-block
load("../data/temp/fitted_models.RData")
```

# Introduction

Are White identifiers those who are percieve economic threats or is it a backlash to a loss of influence over politics? The existing work offers an unclear answer to this question. This note attempts to provide an ombnibus test to determine not only which of these questions has more empirical support but also which among common proxies of economic and political power in the United States appear to be reliable predictors of a White individual's expression of racial identity.

A common source of data and of measures for researchers in the United States is the American National Election Study. I use a penalized regression using data from the 2012, 2016, and 2020 American National Election Studies to examine the degree to which individual measures of economic and political influence explain an individual's uptake of White identity as well as to examine the degree to which changes in the political context may increase or decrease the explanatory power of these measures. I find that proxies reporting concerns about Whites' political influence are stronger explantory factors than proxies of economic influence in the post-Trump era.

# Identity and white political identity

One perspective on what explains racial identity among Whites in the United States are that they are responding to threats to their economic status and power. Realistic Conflict Theory suggests that a racial identity and racial prejudice come as a response to a collective concern among members of a particular racial group about the loss of economic influence and status a group holds relative to other racial groups [@blumer_1958_psr;@bobo_hutchings_1996_asr]. Some evidence suggests that Whites living in areas with shrinking job opportunities respond to this with an increased desire to protect the interests of their racial confederates [@baccini_weymouth_2021_apsr]. Others have experimental evidence supporting this perspective that White identity is motivated by material concerns [@perez_et-al_2022_pp]. Further, @jardina_2019_cup who provides one of the most comprehensive treatments of conceptualizing, measuring, and examining the outcomes of White identity in politics includes indicators of economic status and power in the primary model.

Social identity theory, and its offshoots, would suggest an alternative perspective: that Whites expressing a racial identity are responding to political, cultural, and societal threats. While these threats are a bit more hard to define, there are a number of examples of Whites expressing prejudice towards other racial groups in response to these sorts of challenges to their power. @tesler_2016 demonstrates that Whites heavily racialized American politics in response to Barack Obama's ascention to the White House. Other work demonstrates that the saliency and reprocussions of aligning with competing racial groups drove non-White political candidates to distance themselves from their racial confederates in an attempt to appease Whites [@stephens-dougan_2020]. Some pundits on Fox News suggest that Whites are in direct competition for jobs with immigrants from South and Central America. Despite these claims, some evidence suggests that Whites' hostility directed toward non-White immigrant groups depends on the degree to which they assimilate [@green_staerkle_2013; see also @hainmueller_hopkins_2014]. In terms of attitudes oriented toward the in-group, @dawson_1994_pup argues that Black Americans hold a strong racial identity that supercedes their class-based identity due to the persistent focus on Black individual's race in their personal and political life. For Whites, their racial identity has long been central to their behavior and the way in which they mobilize to shape legal and political institutions to maintain power [@rile-hayward_2013].

# Political versus economic threats

Identity is largely an elusive concept to operationalize. Though there are implicit measures of White identity [see @knowles_peng_2005_jpsp], the common and measure of white identity is a response to the question, "How important is being White to your identity?" This measure captures a significant feature of an identity: the centrality or the importance of that identity to an individual. Though a rather rough measure of something as complex as identity, it has been shown to produce similar results to more complex measures of white identity and holds a significant amount of face validity [@jardina_2019_cup].

Threat is also quite difficult to measure directly. However, we can make a number of reasonable guesses about who may feel economic and political threat. There are two key things that threats affect. Status and power. Status refers to the respectability of the group whereas power refers to the influence that the group has [@craig_phillips_2023_ohpp]. Threat here refers to either a manifestation of economic or political challenges to the current racial hierarchy. I do not distinguish between these two sources of threat as common survey measures are too crude to distinguish between how the threat manifests. Here, I am focusing on what types of non-material threats activate White identity in the context of politics and will, as a shorthand refer to political, cultural, and societal threats as political threats.

A reasonable approach to measuring economic threat is to measure the economic circumstances that White individuals find themselves in. We may expect that Whites who are facing bottom-up economic competition, will be those that experience the strongest desire to identify themselves with their racial group. As the dominant racial group in the United States, Whites that are unemployed, are surrounded by other Whites who are experiencing economic hardship and those who have limited capacity to switch careers are those that may be the most motivated to identify with the dominant racial group to defend Whites' influence and status in society [@bobo_hutchings_1996_asr]. Therefore, measures of employment, finances, family finances, finances of those they know, and educational attainment all seem likely predictors of racial identity for Whites. This fits with measures that @dawson_1994_pup used to measure the influence of economics in Black American's identity. These measures also fit with @lindsay_2023_apr who found that many of these items predicted racial resentment among Whites.

Here I conceptualize of political threat as perceptions that politics and political institutions no longer serves White people. Existing work demonstrates that Whites who express a desire to maintain the racial status quo with Whites on top are more likely to express that Whites are discriminated against at similar levels as Blacks [@marshburn_et-al_2022_gpir]. As Whiteness is often a central component of Whites' conceptualization of American society [see @green_staerkle_2013], changes to "traditional American values" may reflect concern about changes to the influences of White values. Finally, a common finding is that expression of White identity is heavily wrapped up in the perception that racial stratifications are by meritocratic means [@lowery_et-al_2012_jpsp]. As members of a dominant social group, when given information that either threatens one's belief that their merits are earned or that their group operates as a way to maintain inequality, we expect lower levels of expression in identification with that group. The classic items of racial resentment construct capture these ideas in that they revolve around the idea that Blacks are not disadvantaged as the result of systematic oppression by Whites. Though many have often used racial resentment as a measure of prejudice which would be a possible outcome of White identity, some make the compelling argument that it is distinct as a concept and operationalization [@davis_wilson_2022_ucp].

<!--
Social identity theory suggests that individuals form attitudes based on groups they associate with, even if those groups are subjectively defined [see @tajfel_turner_1986]. In the famous Robber's Cave experiment, @sherif_1988 divided a group of adolescent boys into two groups arbitrarily. After a week of spending time with others in the group, the boys came up with group names that defined the two groups and distinguished themselves; one group called themselves the Eagles, and the other group referred to themselves as the Rattlers. With the introduction of competition, these boys reported fierce hostility toward the members of the other group and expressed positive attitudes toward those within the group. 
<!--
This example illustrates several features of social identities. The first feature of social identity theory is that these groups we may associate ourselves with are often subjectively defined [see @huddy_2001]. The second feature is that these subjective definitions of the group make distinctions between the two groups that come to matter to members of each group - some are part of our group, and some are part of that other group. Many cognitive theories suggest that this occurs as a result of the tendency for humans to reduce cognitive load by grouping things together [@huddy_2001]. The third feature is that introducing competition between these groups shapes attitudes toward those within and without your group. The motivator is that groups with higher status often want to keep that status, and individuals in the high-status group are much more likely to support inequality to satisfy that motivation [@burns_kinder_2012;@morrison_et-al_2009]. Fourth, individuals firmly hold these attitudes, which tend to be long-lasting. An additional feature of theories of social identity is that the degree to which one might identify with one group or another is fluid and that whether one identity that one might hold is active or if it is another often depends on the context and the identity's relevance to that context [@huddy_2012].

Racial groups in the U.S. contain a number of these features. As a result, a burgeoning literature in political science that has made this connection takes the racial identity of whites more seriously as one that informs their political attitudes [see @jardina_2019_cup]. Race in America is subjectively defined [see @rile-hayward_2013], and these definitions are deeply embedded and reinforced by our institutions [@lopez_2006]. Whiteness and one's willingness to identify with that group is important [@wong_cho_2005] as it reflects a desire for an association with the history and affordances of being white [@mcdermott_samson_2005]. 
->
Recent work makes the case that we should seriously consider whites' racial identity as an identity [@jardina_2019_cup]. As an identity, much of the literature view these attitudes resulting from whites' racial identity as stemming from economic threat. Some studying white political identity find economic predictors as substantively and statistically significant [see @perez_et-al_2022_pp;@lindsay_2023_apr on their racial attitudes writ large]. Observational evidence supports this claim as well. For example, Trump's messaging about job market competition with immigrants appeared to be successful at attracting white voters [@bunyasi_2019]. After all, the identity literature, more broadly, often focuses on economic threats. Evolutionary psychologists often take the position that identity and the resulting attitudes and behaviors directed toward out-groups are the result of competition over material resources [see @huddy_2001]. Many political scientists use this as an explanation for negative attitudes directed toward immigrants among those who feel insecure about their job prospects [@pardos-prado_xena_2019]. Others argue that deindustrialization and job loss activate whites' concerns about the racial hierarchy and predict their support for Republican challengers that express support for keeping the racial status quo [@baccini_weymouth_2021_apsr].

As much of the literature on whites' racial attitudes rely heavily on social identity theory as a paradigm, a significant portion of it presents the source of these group-based views as responses to economic threat. However, some disagree with this position [see @jardina_2019_cup] and view these attitudes as the result of changes to the racial hierarchy. These other perspectives provide explanations for whites' attitudes and conceptualize the definition of whiteness and white identity as a desire to retain political power.

# Why might white identity be a backlash to loss of institutional privelege?

While a critical component of identity is a feeling of your fate as linked with others in your group [@leach_et-al_2008], this linked fate is the connection between attitudes and material forms of threat but that identity, in and of itself, is better conceptualized as a result of cultural or ideational threat [@donnelly_2020]. Aligning with the contextual account of racial identity for whites, many conceptualize white identity as an expressed desire for a recognition of privilege and status by institutions [@lopez_2006] and a connection with the history that connects whites to their status [@mcdermott_samson_2005]. The prominent account of white political identity and its implications in contemporary American politics argues for a strong connection between white political identity and a desire to protect this racial hierarchy and status. This conceptualization reflects the tendency to report in-group-oriented, not necessarily out-group-oriented, attitudes [@jardina_2019_cup].

Whites' institutional privelege and threats to it are the more potent threats that whites may be less used to and as a result, coalesce as a group to defend. @winter_2008 demonstrates that whites' negative attitudes toward social welfare programs are motivated by perceptions that beneficiaries are Black. In this book length examination of the question, it is not that the people are poor and could use government assistance, it is that those receiving help are not white. Further, the increasing racial polarization is not something that is occuring among whites who are most economically threatened, but the Republican party is whiter with both wealthy and less wealthy whites -- political considerations are causally prior to racial attitudes [@engelhardt_2021_bjps; see also @schaffner_2022_bjps].

An example of a contemporary institutional threat to whites is the election of Barack Obama. Evidence suggests that America did not shift into a post-racial era as some political commentators suspected. In fact, race became an extremely salient issue for whites [@tesler_2016]. White Republicans exhibited higher levels of old fashioned racism [@tesler_2013]. Issues were racialized during Obama's presidency and many Whites exhibited lower levels of support for those policies in response [@tesler_2012;@tesler_2016]. As a result of whites' intense concern with race, non-white political candidates, such as Obama, often had to distance themselves from their racial groups in speeches before white audiences as a way to not activate concerns about being to beholden to their racial group [@stephens-dougan_2020]. The reason for such potent responses to indicators of change for what groups have influence in political instittuions is simply that whites are not used to it. They have spent the entirety of this country's existence defending the idea that whites have higher status in American society regardless of wealth, education, or other forms of status.

My argument is not a new one. Historical accounts outlining the complicated history of what it meant to be "white" from a legal standpoint boils down to: did other whites see you as being good enough to be called white? There are many examples of groups that are now considered white but have not always been so. For example groups from eastern europe were often considered as non-white. While these definitions often correlated with economic status (e.g., German immigrants were often considered white but not Irish immigrants), they often changed in response to concerns about weakening white political influence with other groups growing in political influence such as the Irish in Boston. During these shifts in definitions of whiteness were no longer about class. Instead, it became about what it meant for access and privileges afforded by institutions [@rile-hayward_2013]. In the $19^{th}$ century, privilege came from one's freedom and wealth. Those who were not enslaved but were poor were not considered white and were of a lower strata than whites who owned land. With the abolition of slavery, the expressed (explicit) concern became not about wealth but now about race. Whites could no longer use the classification of "free" and "land-owning" as an excuse to exclude Blacks from full access to America's institutions. This motivated a shift in how the institutions and how whites defined themselves [@rile-hayward_2013]. It became the focus for whites to be a higher status group, regardless of whether or not they were poorer than formerly enslaved Black Americans [@rile-hayward_2013]. In other words, the white American tradition is to hold one's whiteness as a status symbol regardless of other forms of status such as class.

Not one single paper or effort to provide empirical evidence can satisfactorily address such a large and complex question. The causal mechanism here is extremely difficult to pin down. In an effort to address this question in a bite-sized chunk, I first start with examining whether or not correlational evidence aligns with my claim that individual-level economic predictors are weaker than perceptions of political influence. In the next section, I examine whether attitudes expressing concern with economic or political threats are more robust correlates of the degree to which someone feels that being white is important to their identity. The goal is not to test particular causal mechanisms but to examine whether my belief and others' that white identity reflects reactions to political threats more than it does economic.
-->
# Data

To examine whether whites' attitudes expressing concern about political threats are more robust correlates than their perceptions of the economy, I use the 2012, 2016, and 2020 American National Election Studies (ANES). The studies are nationally representative questions that capture several questions that lend themselves to testing the relationship between economic and political concerns while also asking white respondents to report how important being white is to their identity.

Though I am using the more crude measure of White identity and am limited in the predictors I can choose from, a larger concern is that the degree to which these threats may vary as a result of changes to the political landscape. Specifically, in 2008 the United States experienced a massive recession that lead hit the housing market quite hard. In 2012, the economy was still recovering. In 2016, Donald Trump became infamous for an election where he made race and gender extremely salient [@banda_cassese_2022_pb]. Finally, in 2020, Trump had been in office for four years with an improving economy. Each of these three years reflect a different context and may offer the opportunity to observe how economic and political threats operate differrently.

Question wording for each of the items in the following models are included in @tbl-anes-measures which is located in the Supplementary Materials. I include descriptive statistics for the measures in these samples in @tbl-2012-descriptive-stats, @tbl-2016-descriptive-stats, and @tbl-2020-descriptive-stats located in the Supplementary Information.

## Evidence from a penalized regression

The primary contribution of this paper is to examine which of a number of competing measures of different types of threat may predict white identity for individuals. As this requires an ombnibus test, I elect to use a penalized regression. I provide a more detailed discussion of the reasons why I use an estimator with an L2 regularization parameter, often refered to as a LASSO model in the supplementary materials. For a brief comment as to why I use this approach is because approach is designed to take models with a large number of predictors and to shrink the coefficients toward zero. In practice, this often means that predictors that has relatively fewer explanatory power have much smaller coefficients than the variables that do provide significant predictive power.

My measure of white identity is ordinal. Therefore, I apply a ordered logistic link function. Though it creates a more complicated computational task, it is a more appropriate model specification given the available measure and my assumptions about the concept [@liddell_kruschke_2018_jesp].

To account for potential problems stemming from missing data, I impute my data by using multiple imputation with chained equations (MICE) where the particular model I use is a random forest model to allow for more flexibility and to reduce model dependence [@buuren_goothuis-oudshoorn_2011; @marbach_2022_pa].^[I impute 10 datasets and simultaneously fit my Bayesian LASSO on each dataset and pool the results. Uncertainty reflected in my reported credible intervals not only reflects the uncertainty from one model, but my reported uncertainty also reflects the uncertainty generated from the imputation procedure. One drawback with this particular approach is that my normalized split-$\hat{R}$ will likely be greater than the widely recognized 1.01 cutoff that indicates model convergence due to the pooling of my multiply imputed datasets.] I report the results of the model fitted using listwise deletion in the Supplementary Information.

## Results

@fig-2012-2016-2020-ordinal-lasso presents the estimates of my fitted ordinal models with the LASSO shrinkage prior.^[The full tables of results are included in the Supplementary Information.] These models are fit on 6 chains with 2000 iterations each. To examine the evidence of whether a particular predictor "matters" for predicting white political identity, I examine the estimated credible intervals, which report the probability that the true estimate falls within the estimated range. I construct these credible intervals by reporting estimates between the 2.5% and 97.5% quantiles. This means that for a given credible interval, there is a two-tailed 95% probability that the true value falls within that range. For values outside the range of the credible interval, it is relatively implausible that the given value is the true value.

```{r}
#| label: fig-2012-2016-2020-ordinal-lasso
#| layout-ncol: 2
#| fig-cap: Economic factors are weak predictors of white identity
#| fig-subcap:
#|   - Economic
#|   - Political
# Graphs
  #* Tidyfy posterior draws
    #** 2012
lasso[['2012_ordinal_tidy']] <- parameters(
  lasso[['2012_ordinal']]
  , conf.int = 0.95
  , verbose = FALSE
) |> # grab the posterior draws and construct CI's at the 95% level
  standardize_names(style = 'broom') |> # standardize the coefficient names
  data.frame() |> # convert it to a data.frame
  mutate(model = '2012') |> # add a column called model with the value of 2012
  filter(!is.na(term)) # get rid of any rows that are na
    #** 2016
lasso[['2016_ordinal_tidy']] <- parameters(
  lasso[['2016_ordinal']]
  , conf.int = 0.95
  , verbose = FALSE
) |> # grab the posterior draws and construct CI's at the 95% level
  standardize_names(style = 'broom') |> # standardize coefficient names
  data.frame() |> # convert it to a data.frame
  mutate(model = '2016') |> # add a column called model with the value of 2016
  filter(!is.na(term)) # get rid of any rows that are na
    #** 2020
lasso[['2020_ordinal_tidy']] <- parameters(
  lasso[['2020_ordinal']]
  , conf.int = 0.95
  , verbose = FALSE
) |> # grab the posterior drawsw and construct CI's at the 95% level
  standardize_names(style = 'broom') |> # standardize coefficient names
  data.frame() |> # convert it to a data.frame
  mutate(model = '2020') |> # add a column called model with the value of 2020
  filter(!is.na(term)) # get rid of any rows that are na

lasso[['ordinal_tidy']] <- rbind(
  lasso[['2012_ordinal_tidy']]
  , lasso[['2016_ordinal_tidy']]
  , lasso[['2020_ordinal_tidy']]
) |> # append the three models' results together
  mutate(
    term = case_when(
      term == 'b_rboff' ~ 'Retrospective better off'
      , term == 'b_epast' ~ 'Economy better'
      , term == 'b_unpast' ~ 'Unemployment better'
      , term == 'b_edu' ~ 'Education'
      , term == 'b_income' ~ 'Income'
      , term == 'b_losejob' ~ 'Worried about losing job'
      , term == 'b_ecfamily' ~ 'Worried about family finances'
      , term == 'b_ecjob' ~ 'Know someone lost job'
      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
      , term == 'b_govbias' ~ 'Gov. biased against whites'
      , term == 'b_inflwhite' ~ 'whites influence politics'
      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
      , term == 'b_raceResent' ~ 'Racial resentment'
      , term == 'b_pid' ~ 'Party ID'
      , term == 'b_female' ~ 'Female'
    ), # rename the parameters
    term = factor(
      term
      , levels = c(
        'Retrospective better off'
        , 'Economy better'
        , 'Unemployment better'
        , 'Education'
        , 'Income'
        , 'Worried about losing job'
        , 'Worried about family finances'
        , 'Know someone lost job'
        , 'New lifestyles breaking society'
        , 'Gov. biased against whites'
        , 'whites influence politics'
        , 'whites are discriminated against'
        , 'Racial resentment'
        , 'Party ID'
        , 'Female'
      )
    ) # convert it to a factor variable
  ) |> 
  filter(!is.na(term)) # get rid of na rows
    #** make a data.frame of results for the econ predictors
lasso[["ordinal_tidy_econ"]] <- lasso[["ordinal_tidy"]] |>
  filter(
    term == "Retrospective better off" | term == "Economy better" | 
    term == "Unemployment better" | term == "Education" | term == "Income" | 
    term == "Worried about losing job" | term == "Wrroeid about family finances" |
    term == "Know someone lost job"
  )
    #** make a data.frame of results for the political predictors
lasso[["ordinal_tidy_poli"]] <- lasso[["ordinal_tidy"]] |>
  filter(
    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
    term == "whites influence politics" | term == "whites are discriminated against" |
    term == "Racial resentment"
  )
  #* plot for econ predictors
econ <- ggplot(
  data = lasso[['ordinal_tidy_econ']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
  #* plot for political predictors
poli <- ggplot(
  data = lasso[['ordinal_tidy_poli']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))

# Print plots
econ
poli
```

Examining the figure, we see that none of the proxies for economic threat consistently correlate with White political identity; however there is at least one proxy of political threat that does. For economic threats, we see that when the economy was recovering from the recession and Obama was in Office, those who perceived that the economy was better off retrospectively and with a lower reported level of education were correlated with expressing a higher White identity. However, when the economy was better and Trump was in office, we see that those less worried about losing a job were reported a higher white identity.

For political threats, we see that in both instances where Obama was in the White House that those who reported being concerned with "new lifestyles" as well as those with a higher score on the racial resentment measures were associated with having higher white political identities. When Trump was in the White House, we see that those who felt that the government was biased against Whites, that Whites were discriminated against, and who scored higher on the racial resentment measures all reported higher levels of White political identity.

An additional question worth examining is that economic or political threat may appear differently for partisans and for female respondents. Recent evidence suggests that racial attitudes are guided by partisanship [@engelhardt_2021_bjps]. Specifically, some evidence suggests that White Democrats are much more sensitive to their racial identity than White Republicans due to the higher amount of racial heterogeneity within the Democratic party. This may mean for White Democrats that they have more exposure to intergroup competition for power within the party (such as the choice to support Hillary Clinton versus Barack Obama) and may have higher White identification as a result. The second way that these threats may matter more or less for Whites is on the dimension of sex. Gender-based differences in economic influence and status remains quite high. We therefore may expect that female individuals are more sensitive to economic threats than their male counterparts. In addition to the ombnibus model, I also include two additional models where I split the samples along the dimension of partisanship (including leaners in each) and on the dimension of sex.

```{r}
#| label: fig-2012-2016-2020-female-male-lasso
#| layout-ncol: 2
#| fig-cap: Economic factors are weak predictors of white identity
#| fig-subcap:
#|   - Female, Economic
#|   - Female, Political
#|   - Male, Economic
#|   - Female, Economicß
# Graphs
  #* Tidyfy posterior draws
    #** 2012
lasso[["2012_female_tidy"]] <- parameters(
  lasso[["2012_female"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2012") |>
  filter(!is.na(term))
lasso[["2012_male_tidy"]] <- parameters(
  lasso[["2012_male"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2012") |>
  filter(!is.na(term))
    #** 2016
  lasso[["2016_female_tidy"]] <- parameters(
  lasso[["2016_female"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2016") |>
  filter(!is.na(term))
lasso[["2016_male_tidy"]] <- parameters(
  lasso[["2016_male"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2016") |>
  filter(!is.na(term))
    #** 2020
lasso[["2020_female_tidy"]] <- parameters(
  lasso[["2020_female"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2020") |>
  filter(!is.na(term))
lasso[["2020_male_tidy"]] <- parameters(
  lasso[["2020_male"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2020") |>
  filter(!is.na(term))

lasso[['female_tidy']] <- rbind(
  lasso[['2012_female_tidy']]
  , lasso[['2016_female_tidy']]
  , lasso[['2020_female_tidy']]
) |> # append the three models' results together
  mutate(
    term = case_when(
      term == 'b_rboff' ~ 'Retrospective better off'
      , term == 'b_epast' ~ 'Economy better'
      , term == 'b_unpast' ~ 'Unemployment better'
      , term == 'b_edu' ~ 'Education'
      , term == 'b_income' ~ 'Income'
      , term == 'b_losejob' ~ 'Worried about losing job'
      , term == 'b_ecfamily' ~ 'Worried about family finances'
      , term == 'b_ecjob' ~ 'Know someone lost job'
      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
      , term == 'b_govbias' ~ 'Gov. biased against whites'
      , term == 'b_inflwhite' ~ 'whites influence politics'
      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
      , term == 'b_raceResent' ~ 'Racial resentment'
      , term == 'b_pid' ~ 'Party ID'
    ), # rename the parameters
    term = factor(
      term
      , levels = c(
        'Retrospective better off'
        , 'Economy better'
        , 'Unemployment better'
        , 'Education'
        , 'Income'
        , 'Worried about losing job'
        , 'Worried about family finances'
        , 'Know someone lost job'
        , 'New lifestyles breaking society'
        , 'Gov. biased against whites'
        , 'whites influence politics'
        , 'whites are discriminated against'
        , 'Racial resentment'
        , 'Party ID'
      )
    ) # convert it to a factor variable
  ) |> 
  filter(!is.na(term)) # get rid of na rows
lasso[['male_tidy']] <- rbind(
  lasso[['2012_male_tidy']]
  , lasso[['2016_male_tidy']]
  , lasso[['2020_male_tidy']]
) |> # append the three models' results together
  mutate(
    term = case_when(
      term == 'b_rboff' ~ 'Retrospective better off'
      , term == 'b_epast' ~ 'Economy better'
      , term == 'b_unpast' ~ 'Unemployment better'
      , term == 'b_edu' ~ 'Education'
      , term == 'b_income' ~ 'Income'
      , term == 'b_losejob' ~ 'Worried about losing job'
      , term == 'b_ecfamily' ~ 'Worried about family finances'
      , term == 'b_ecjob' ~ 'Know someone lost job'
      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
      , term == 'b_govbias' ~ 'Gov. biased against whites'
      , term == 'b_inflwhite' ~ 'whites influence politics'
      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
      , term == 'b_raceResent' ~ 'Racial resentment'
      , term == 'b_pid' ~ 'Party ID'
    ), # rename the parameters
    term = factor(
      term
      , levels = c(
        'Retrospective better off'
        , 'Economy better'
        , 'Unemployment better'
        , 'Education'
        , 'Income'
        , 'Worried about losing job'
        , 'Worried about family finances'
        , 'Know someone lost job'
        , 'New lifestyles breaking society'
        , 'Gov. biased against whites'
        , 'whites influence politics'
        , 'whites are discriminated against'
        , 'Racial resentment'
        , 'Party ID'
      )
    ) # convert it to a factor variable
  ) |> 
  filter(!is.na(term)) # get rid of na rows
    #** make a data.frame of results for the econ predictors
lasso[["female_tidy_econ"]] <- lasso[["female_tidy"]] |>
  filter(
    term == "Retrospective better off" | term == "Economy better" | 
    term == "Unemployment better" | term == "Education" | term == "Income" | 
    term == "Worried about losing job" | term == "Wrroeid about family finances" |
    term == "Know someone lost job"
  )
    #** make a data.frame of results for the political predictors
lasso[["female_tidy_poli"]] <- lasso[["female_tidy"]] |>
  filter(
    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
    term == "whites influence politics" | term == "whites are discriminated against" |
    term == "Racial resentment"
  )
lasso[["male_tidy_econ"]] <- lasso[["male_tidy"]] |>
  filter(
    term == "Retrospective better off" | term == "Economy better" | 
    term == "Unemployment better" | term == "Education" | term == "Income" | 
    term == "Worried about losing job" | term == "Wrroeid about family finances" |
    term == "Know someone lost job"
  )
    #** make a data.frame of results for the political predictors
lasso[["male_tidy_poli"]] <- lasso[["male_tidy"]] |>
  filter(
    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
    term == "whites influence politics" | term == "whites are discriminated against" |
    term == "Racial resentment"
  )
female_econ <- ggplot(
  data = lasso[['female_tidy_econ']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
  #* plot for political predictors
female_poli <- ggplot(
  data = lasso[['female_tidy_poli']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
male_econ <- ggplot(
  data = lasso[['male_tidy_econ']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
  #* plot for political predictors
male_poli <- ggplot(
  data = lasso[['male_tidy_poli']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))

female_econ
female_poli
male_econ
male_poli
```

```{r}
#| label: fig-2012-2016-2020-pid-lasso
#| layout-ncol: 2
#| fig-cap: Economic factors are weak predictors of white identity
#| fig-subcap:
#|   - Republican, Economic
#|   - Republican, Political
#|   - Democrat, Economic
#|   - Democrat, Economic
# Graphs
  #* Tidyfy posterior draws
    #** 2012
lasso[["2012_republican_tidy"]] <- parameters(
  lasso[["2012_republican"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2012") |>
  filter(!is.na(term))
lasso[["2012_democrat_tidy"]] <- parameters(
  lasso[["2012_democrat"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2012") |>
  filter(!is.na(term))
    #** 2016
  lasso[["2016_republican_tidy"]] <- parameters(
  lasso[["2016_republican"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2016") |>
  filter(!is.na(term))
lasso[["2016_democrat_tidy"]] <- parameters(
  lasso[["2016_democrat"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2016") |>
  filter(!is.na(term))
    #** 2020
lasso[["2020_republican_tidy"]] <- parameters(
  lasso[["2020_republican"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2020") |>
  filter(!is.na(term))
lasso[["2020_democrat_tidy"]] <- parameters(
  lasso[["2020_democrat"]]
  , conf.int = 0.95
  , verbose = FALSE
) |>
  standardize_names(style = "broom") |>
  data.frame() |>
  mutate(model = "2020") |>
  filter(!is.na(term))

lasso[['republican_tidy']] <- rbind(
  lasso[['2012_republican_tidy']]
  , lasso[['2016_republican_tidy']]
  , lasso[['2020_republican_tidy']]
) |> # append the three models' results together
  mutate(
    term = case_when(
      term == 'b_rboff' ~ 'Retrospective better off'
      , term == 'b_epast' ~ 'Economy better'
      , term == 'b_unpast' ~ 'Unemployment better'
      , term == 'b_edu' ~ 'Education'
      , term == 'b_income' ~ 'Income'
      , term == 'b_losejob' ~ 'Worried about losing job'
      , term == 'b_ecfamily' ~ 'Worried about family finances'
      , term == 'b_ecjob' ~ 'Know someone lost job'
      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
      , term == 'b_govbias' ~ 'Gov. biased against whites'
      , term == 'b_inflwhite' ~ 'whites influence politics'
      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
      , term == 'b_raceResent' ~ 'Racial resentment'
      , term == 'b_female' ~ 'Female'
    ), # rename the parameters
    term = factor(
      term
      , levels = c(
        'Retrospective better off'
        , 'Economy better'
        , 'Unemployment better'
        , 'Education'
        , 'Income'
        , 'Worried about losing job'
        , 'Worried about family finances'
        , 'Know someone lost job'
        , 'New lifestyles breaking society'
        , 'Gov. biased against whites'
        , 'whites influence politics'
        , 'whites are discriminated against'
        , 'Racial resentment'
        , 'Female'
      )
    ) # convert it to a factor variable
  ) |> 
  filter(!is.na(term)) # get rid of na rows
lasso[['democrat_tidy']] <- rbind(
  lasso[['2012_democrat_tidy']]
  , lasso[['2016_democrat_tidy']]
  , lasso[['2020_democrat_tidy']]
) |> # append the three models' results together
  mutate(
    term = case_when(
      term == 'b_rboff' ~ 'Retrospective better off'
      , term == 'b_epast' ~ 'Economy better'
      , term == 'b_unpast' ~ 'Unemployment better'
      , term == 'b_edu' ~ 'Education'
      , term == 'b_income' ~ 'Income'
      , term == 'b_losejob' ~ 'Worried about losing job'
      , term == 'b_ecfamily' ~ 'Worried about family finances'
      , term == 'b_ecjob' ~ 'Know someone lost job'
      , term == 'b_tradbreak' ~ 'New lifestyles breaking society'
      , term == 'b_govbias' ~ 'Gov. biased against whites'
      , term == 'b_inflwhite' ~ 'whites influence politics'
      , term == 'b_whitediscrim' ~ 'whites are discriminated against'
      , term == 'b_raceResent' ~ 'Racial resentment'
      , term == 'b_female' ~ 'Female'
    ), # rename the parameters
    term = factor(
      term
      , levels = c(
        'Retrospective better off'
        , 'Economy better'
        , 'Unemployment better'
        , 'Education'
        , 'Income'
        , 'Worried about losing job'
        , 'Worried about family finances'
        , 'Know someone lost job'
        , 'New lifestyles breaking society'
        , 'Gov. biased against whites'
        , 'whites influence politics'
        , 'whites are discriminated against'
        , 'Racial resentment'
        , 'Female'
      )
    ) # convert it to a factor variable
  ) |> 
  filter(!is.na(term)) # get rid of na rows
    #** make a data.frame of results for the econ predictors
lasso[["republican_tidy_econ"]] <- lasso[["republican_tidy"]] |>
  filter(
    term == "Retrospective better off" | term == "Economy better" | 
    term == "Unemployment better" | term == "Education" | term == "Income" | 
    term == "Worried about losing job" | term == "Wrroeid about family finances" |
    term == "Know someone lost job"
  )
    #** make a data.frame of results for the political predictors
lasso[["republican_tidy_poli"]] <- lasso[["republican_tidy"]] |>
  filter(
    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
    term == "whites influence politics" | term == "whites are discriminated against" |
    term == "Racial resentment"
  )
lasso[["democrat_tidy_econ"]] <- lasso[["democrat_tidy"]] |>
  filter(
    term == "Retrospective better off" | term == "Economy better" | 
    term == "Unemployment better" | term == "Education" | term == "Income" | 
    term == "Worried about losing job" | term == "Wrroeid about family finances" |
    term == "Know someone lost job"
  )
    #** make a data.frame of results for the political predictors
lasso[["democrat_tidy_poli"]] <- lasso[["democrat_tidy"]] |>
  filter(
    term == "New lifestyles breaking society" | term == "Gov. biased against whites" |
    term == "whites influence politics" | term == "whites are discriminated against" |
    term == "Racial resentment"
  )
republican_econ <- ggplot(
  data = lasso[['republican_tidy_econ']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
  #* plot for political predictors
republican_poli <- ggplot(
  data = lasso[['republican_tidy_poli']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
democrat_econ <- ggplot(
  data = lasso[['democrat_tidy_econ']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))
  #* plot for political predictors
democrat_poli <- ggplot(
  data = lasso[['democrat_tidy_poli']]
  , aes(
    y = term
    , x = estimate
    , xmin = conf.low
    , xmax = conf.high
    , shape = factor(model)
  )
) +
  geom_pointinterval(
    position = position_dodge(width = 0.6)
    , size = 4
) +
  geom_vline(
    xintercept = 0
    , linetype = 2
) +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  labs(
    shape = 'Model'
    , x = 'Estimate'
    , y = ''
    , caption = 'Data Source: 2012, 2016, 2020 ANES.\n Dots are median draw from posterior distribution.\n Bars represent 95% credible intervals.'
) + 
  theme(text=element_text(size=12, family = 'sans'))

republican_econ
republican_poli
democrat_econ
democrat_poli
```

We see in both @fig-2012-2016-2020-female-male-lasso and @fig-2012-2016-2020-pid-lasso that indicators of economic and political threat do not have much substantial variation between male and female respondents and Republican and Democrat respondents. The full table of results for the model are included in @tbl-2012-gender-lasso, @tbl-2016-gender-lasso, @tbl-2020-gender-lasso, @tbl-2012-pid-lasso, @tbl-2016-pid-lasso, and @tbl-2020-pid-lasso in the Supplimentary Materials.

# Discussion and conclusions

Overall, these results suggest that common proxies of economic threat are at least inconsistent measures of White identity. Even in a time where the economy is recovering from a collapse, more proxies of political threat seemed to predict White identity. When a White president well-known for his penchance for racial dog-whistles takes office, we see support for the idea that White grievance politics was at a high -- high those who felt that political institutions and that the direction of society was turning away from the racial status quo were those that reported their racial identity as more central to who they are.

There are a few important points that are illustrated here. First, suggestions that White political identity is something occuring for those who are most economically threatened appear to have weak theoretical and empirical backing when directly pitted against other forms of threat. Second, that measures of cultural or societal threat are important variables to include if one is using white identity as a variable in a model.

These findings provide a direction for scholars to take when examining the factors driving Whites to take on a white political identity. While this particular analysis is not causal, it provides a basis for those wondering whether, when put together, economic or political threats are more fruitful to engender a rise in white political identity.

{{< pagebreak >}}

# References

::: {#refs}
:::

{{< pagebreak >}}

# Supplementary Information {.appendix}

## Measures {.appendix}

\blandscape
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
|                                              | Abbreviated question                                       | ANES Coding                                                                                                                                                  | Author coding                                                                                                                                                             | Years            |
+==============================================+============================================================+==============================================================================================================================================================+===========================================================================================================================================================================+==================+
| white identity                               | How important is being white to your identity              | 1 = Extremely important; 2 = Very important; 3 = Moderately important; 4 = A little important; 5 = Not at all important; \<1 Missing, not asked, etc         | 1 = Not at all important; 2 = A little important; 3 = Moderately important; 4 = Very important; 5 = Extremely important; NA = Missing, not asked, etc                     | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Economics**                                |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Retrospective better off                     | How much better worse off than 1 year ago                  | 1 = Much better; 2 = Somewhat better; 3 = The same; 4 = Somewhat worse; 5 = Much worse, \< 1 Missing, not asked, etc                                         | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Economy better                               | U.S. economy better worse off than 1 year ago              | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                            | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Unemployment better                          | Unemployment better or worse than 1 year ago               | 1 = Much better; 2 = somewhat better; 3 = Stayed about the same, 4 = Somewhat worse; 5 = Much worse; \< 1 Missing, not asked, etc                            | -2 = Much worse; -1 = Somewhat worse; 0 = The same; 1 = Somewhat better; 2 = Much better; NA = Missing, not asked, etc                                                    | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education                                    | Educational attainment                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; \< 1 Missing, not asked, etc                                     | 1 = \< than High school; 2 = High school; 3 = Some post-High school; 4 = Bachelor's degree; NA = Missing, not asked, etc                                                  | 2012, 2020       |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Education (2016)                             | Educational attainment                                     | See ANES codebook                                                                                                                                            | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2016             |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Income                                       | Total family income                                        | See ANES codebook; \< Missing, not asked, etc                                                                                                                | See ANES codebook; NA = MIssing, not asked, etc                                                                                                                           | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about losing job                     | How worried about losing job in near future                | 1 = Not at all; 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; \< 1 Missing, not asked, etc                                                          | -2 = Not at all; -1 = A little; 0 = Moderately; 1 = Very; 2 = Extremely                                                                                                   | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Immigrants take jobs                         | How likely immigration will take away jobs                 | 1 = Extremely; 2 = Very; 3 = Somewhat; 4 = Not at all; \< 1 Missing, not asked, etc                                                                          | 1 = Not at all; 2 = Somewhat; 3 = Very; 4 = Extremely; NA = Missing, not asked, etc                                                                                       | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Worried about family finances                | Worry about family financial situation                     | 1 = Extremely; 2 = Very; 3 = Moderately; 4 = A little; 5 = Not at all; \< 1 Missing, not asked, etc                                                          | 1 = Not at all, 2 = A little; 3 = Moderately; 4 = Very; 5 = Extremely; NA = Missing, not asked, etc                                                                       | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Know someone lost job                        | Know someone who lost job                                  | 1 = Someone lost job; 2 = No one lost job; \< 1 Missing, not asked, etc                                                                                      | 0 = No one lost job; 1 = Someone lost job; NA = Missing, not asked, etc                                                                                                   | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Political**                                |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| New lifestyles breaking society              | Newer lifestyles breaking down society                     | 1 = Agree strongly; 2 = Agree somewhat; 3 = Neither; 4 = Disagree somewhat; 5 = Disagree strongly; \< 1 Missing, not asked, etc                              | -2 = Disagree strongly; -1 = Disagree somewhat; 0 = Neither, 1 = Agree somewhat; 2 = Agree strongly                                                                       | 2012, 2016       |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Government biased against whites             | Does the Administration favor Blacks or whites (2012 only) | 1 = Favors whites; 2 = Favors Blacks; 3 = Treats both the same; \< 1 Missing, not asked, etc                                                                 | -1 = Favors whites; 0 = Treats both the same; 1 = Favors Blacks; NA = Missing, not asked, etc                                                                             | 2012, 2016, 2020 |
|                                              |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
|                                              | Does the Federal Gov treat Blacks or whites Better         |                                                                                                                                                              |                                                                                                                                                                           |                  |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| whites influence politics                    | How much influence do whites have on U.S. politics         | 1 = Too much; 2 = Just about right; 3 = Too little; \< 1 Missing, not asked, etc                                                                             | -1 = Too little; 0 = Just about right; 1 = Too much; NA = Missing, not asked, etc                                                                                         | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| whites are discriminated against             | Discrimination in U.S. against whites                      | 1 = A great deal; 2 = A lot; 3 = A moderate amount; 4 = A little; 5 = None at all; \< 1 Missing, not asked, etc                                              | 1 = None at all; 2 = A little; 3 = A moderate amount; 4 = A lot; 5 = A great deal; NA = Missing, not asked, etc                                                           | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| **Controls**                                 |                                                            |                                                                                                                                                              |                                                                                                                                                                           |                  |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Racial Resentment                            | Standard, 4-item, battery                                  | 5-point Likert scale                                                                                                                                         | Avg of four questions; NA = Missing, not asked, etc                                                                                                                       | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Party ID                                     | Standard battery                                           | 1 = Strong  Dem; 2 = De; 3 = Leans Dem; 4 = Independent; 5 = Leans Rep; 6 = Rep; 7 = Strong Rep; \< 1 Missing, not asked, etc                                | 1 = Strong Democrat; 2 = Democrat; 3 = Leans Democrat; 4 = Independent/Neither; 5 = Leans Republican; 6 = Republican; 7 = Strong Republican; \< 1 Missing, not asked, etc | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
| Female                                       | Sex of respondent                                          | 1 = Male, 2 = Female; \< 1 Missing, not asked, etc                                                                                                           | 0 = Male, 1 = Female; NA = Missing, not asked, etc                                                                                                                        | 2012, 2016, 2020 |
+----------------------------------------------+------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------+
: ANES Measures {#tbl-anes-measures tbl-colwidths="[20,20,25,25,10]"}

\elandscape
{{< pagebreak >}}

## Descriptive statistics {.appendix}

```{r}
#| label: loading-data-block
#create empty list object to hold datasets in
data <- list()
# load the cleaned data
    #* 2012
data[["anes_2012"]] <- read.csv("../data/clean/anes_2012_updated.csv")

data[["anes_2012"]] <- data[["anes_2012"]][
  , c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
    )
]
data[["anes_2012"]] <- setnames(
  data[["anes_2012"]]
  , old = c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
  )
  , new = c(
    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against whites", "Whites influence politics"
    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
  ))

    #* 2016
data[["anes_2016"]] <- read.csv("../data/clean/anes_2016_updated.csv") 
data[["anes_2016"]] <- data[["anes_2016"]][
  , c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
    )
]
data[["anes_2016"]] <- setnames(
  data[["anes_2016"]]
  , old = c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "tradbreak", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
  )
  , new = c(
    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
    , "Worried about family finances", "Know someone lost job", "New lifestyles breaking tradition", "Government biased against whites", "Whites influence politics"
    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
  ))
    #* 2020
data[["anes_2020"]] <- read.csv("../data/clean/anes_2020_updated.csv") 
data[["anes_2020"]] <- data[["anes_2020"]][
  , c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
    )
]
data[["anes_2020"]] <- setnames(
  data[["anes_2020"]]
  , old = c(
    "wid", "rboff", "epast", "unpast", "edu", "income", "losejob"
    , "ecfamily", "ecjob", "govbias", "inflwhite"
    , "whitediscrim", "raceResent", "pid", "female"
  )
  , new = c(
    "White identity", "Retrospective better off", "Economy better", "Unemployment better", "Education", "Income", "Worried about losing job"
    , "Worried about family finances", "Know someone lost job", "Government biased against whites", "Whites influence politics"
    , "Whites discrimminated against", "Racial resentment", "Party ID", "Female"
  ))
```

```{r}
#| label: tbl-2012-descriptive-stats
#| tbl-cap: 2012 ANES Descriptive Statistics
datasummary_skim(
  data[["anes_2012"]]
  , histogram = FALSE
  , notes = c(
    "NAs are included in Unique column."
    , "Data source: 2012 American National Election Study, unweighted."
  )
  , output = "latex"
)
```

```{r}
#| label: tbl-2016-descriptive-stats
#| tbl-cap: 2016 ANES Descriptive Statistics

datasummary_skim(
  data[["anes_2016"]]
  , histogram = FALSE
  , notes = c(
    "NAs are included in Unique column."
    , "Data source: 2016 American National Election Study, unweighted."
  )
  , output = "latex"
)
```

```{r}
#| label: tbl-2020-descriptive-stats
#| tbl-cap: 2020 ANES Descriptive Statistics

datasummary_skim(
  data[["anes_2020"]]
  , histogram = FALSE
  , notes = c(
    "NAs are included in Unique column."
    , "Data source: 2020 American National Election Study, unweighted."
  )
  , output = "latex"
)
```

{{< pagebreak >}}

## Model{.appendix}

The literature surveyed above suggests that there are a small number of variables that do predict white political identity, and there are many others that have no effect. Imagining an assumed hypothesis about the distribution of coefficients, this implies that the L2 regularization parameter equals 0. This is because the L2 regularization parameter shrinks coefficients asymptotically toward zero. The L1 regularization parameter, however, does allow for a shrinkage parameter to pull coefficients to 0. Given this, it suggests that the more appropriate model to capture this debate would be the LASSO penalized regression.

Rather than use the classical LASSO with Leave One Out Cross Validation (LOOCV), I take advantage of a type of Markov Chain Monte Carlos (MCMC) called a Hamiltonian Monte Carlo for my LASSO due to a number of benefits outlined by @van-erp_et-al_2019_jmp. As the literature leads me to hold *apriori* expectations about the distribution of predictors that meaningfully effect white political identity, I specify a Laplace distribution as my prior density function. I suspect that there are a large number of included predictors that have null effects and a relatively smaller proportion of predictors that *do* have an effect but am relatively unsure about how large their coefficients might be. This characterization of my prior fits with that of an L2 Shrinkage parameter specified in a LASSO regression. As I am not using LOOCV, I am able to interpret the results of the model through my credible intervals.

### Full tables of results {.appendix}

```{r}
#| label: model-table-setup-block
    #* Set up non-coefficient related information from models
gm <- list(
    list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)
)

cm <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov. biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')

cm_gender <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov. biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_pid' = 'Party ID', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')

cm_pid <- c('b_rboff' = 'Retrospective better off', 'b_epast' = 'Economy better', 'b_unpast' = 'Unemployment better', 'b_edu' = 'Education', 'b_income' = 'Income', 'b_losejob' = 'Worried about losing job', 'b_ecfamily' = 'Worried about family finances', 'b_ecjob' = 'Know someone lost job', 'b_tradbreak' = 'New lifestyles breaking society', 'b_govbias' = 'Gov. biased against whites', 'b_inflwhite' = 'whites influence politics', 'b_whitediscrim' = 'whites are discriminated against', 'b_raceResent' = 'Racial resentment', 'b_female' = 'Female', 'b_Intercept[1]' = 'Threshold 1', 'b_Intercept[2]' = 'Threshold 2', 'b_Intercept[3]' = 'Threshold 3', 'b_Intercept[4]' = 'Threshold 4')
```

```{r}
#| label: tbl-2012-lasso-simple
#| tbl-cap: Predictors of white Identity in 2012
model_2012 <- list(
  "2012 LWD" = lasso[["2012_simple"]]
  , "2012" = lasso[["2012_ordinal"]]
)
rows <- tribble(~term, ~`2012 LWD`, ~`2012`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 32)

modelsummary(model_2012, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))

```

```{r}
#| label: tbl-2012-gender-lasso
#| tbl-cap: Predictors of Whit Identity in 2012 conditional on gender

model_2012_gender <- list(
  "2012 Female" = lasso[["2012_female"]]
  , "2012 Male" = lasso[["2012_male"]]
)
rows <- tribble(~term, ~`2012 Female`, ~`2012 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2012_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
```

```{r}
#| label: tbl-2012-pid-lasso
#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship

model_2012_pid <- list(
  "2012 Republican" = lasso[["2012_republican"]]
  , "2012 Democrat" = lasso[["2012_democrat"]]
)
rows <- tribble(~term, ~`2012 Republican`, ~`2012 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2012_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2012 American National Election Study.'))
```

```{r}
#| label: tbl-2016-lasso-simple
#| tbl-cap: Predictors of white identity in 2016
model_2016 <- list(
  "2016 LWD" = lasso[["2016_simple"]]
  , "2016" = lasso[["2016_ordinal"]]
)
rows <- tribble(~term, ~`2016 LWD`, ~`2016`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 32)

modelsummary(model_2016, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
```

```{r}
#| label: tbl-2016-gender-lasso
#| tbl-cap: Predictors of Whit Identity in 2016 conditional on gender

model_2016_gender <- list(
  "2016 Female" = lasso[["2016_female"]]
  , "2016 Male" = lasso[["2016_male"]]
)
rows <- tribble(~term, ~`2016 Female`, ~`2016 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2016_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
```

```{r}
#| label: tbl-2016-pid-lasso
#| tbl-cap: Predictors of Whit Identity in 2016 conditional on partisanship

model_2016_pid <- list(
  "2016 Republican" = lasso[["2016_republican"]]
  , "2016 Democrat" = lasso[["2016_democrat"]]
)
rows <- tribble(~term, ~`2016 Republican`, ~`2016 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2016_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2016 American National Election Study.'))
```
```{r}
#| label: tbl-2020-lasso-simple
#| tbl-cap: Predictors of white identity in 2020
model_2020 <- list(
  "2020 LWD" = lasso[["2020_simple"]]
  , "2020" = lasso[["2020_ordinal"]]
)
rows <- tribble(~term, ~`2020 LWD`, ~`2020`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 25, 30)

modelsummary(model_2020, coef_map = cm, statistic = 'conf.int', gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
```

```{r}
#| label: tbl-2020-gender-lasso
#| tbl-cap: Predictors of Whit Identity in 2020 conditional on gender

model_2020_gender <- list(
  "2020 Female" = lasso[["2020_female"]]
  , "2020 Male" = lasso[["2020_male"]]
)
rows <- tribble(~term, ~`2020 Female`, ~`2020 Male`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2020_gender, coef_map = cm_gender, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
```

```{r}
#| label: tbl-2020-pid-lasso
#| tbl-cap: Predictors of Whit Identity in 2012 conditional on partisanship

model_2020_pid <- list(
  "2020 Republican" = lasso[["2020_republican"]]
  , "2020 Democrat" = lasso[["2020_democrat"]]
)
rows <- tribble(~term, ~`2020 Republican`, ~`2020 Democrat`, 'Economic', '', '', 'Non-material', '', '', 'Demographics', '', '', 'Thresholds', '', '')
attr(rows, 'position') <- c(1, 18, 27, 30)

modelsummary(model_2020_pid, coef_map = cm_pid, statistic = "conf.int", gof_map = NA, add_rows = rows, notes = c('Median estimate from fitted model with 6 chains and 2000 iterations.', '95 percent credible intervals in brackets.', 'Data source: 2020 American National Election Study.'))
```